### 2025-01-29 Chapters/Chapter3.tex
Updated tables and appendices

This project used three main data sources, which need to be queried, filtered and combined to prepare the data for use in the models. When working with hundred of thousands of rows, the efficiency of the code is very important. Iterating through those rows might be necessary at times but will increase the time exponentially as compared to using vectorizing methods were possible. The three data sources were all in a different format. Measurement data from IMO was in text files, elevation data was in GeoTiff and reanalysis data from CARRA was in a GRIB format. To use the data to train, these three data sources needed to be combined into one file. This was done based on the measurement data from the IMO. A limit was set on the average wind speed and it was used to select measurement points. Along with the average wind speed having to be above a certain limit, to ensure that the same weather for the same location is not duplicated. The data from IMO was supplied for 10 minute increments, while CARRA data is in 3 hour intervals. This means that to use the CARRA data to predict the measured values from IMO, temporal interpolation would need to be done. Along with the temporal interpolation, note that the CARRA data is given in a rectangular grid where the distance between each point is around 2.5 km while the the information from the IMO is given at specific locations. The elevation information was given by a 20 by 20 rectangular grid that covers Iceland. When combining these data sources an interpolation method needs to be decided upon. Here linear interpolation was applied, both temporally and spatially. The choice of interpolation method, although potentially impactful on the results, was not specifically addressed in this study.

@@

Once CARRA data has been merged with AWS data, using station and time columns, then this combined file needs to checked for nails. This is done by using the hourly data (which is supposedly error free). As the data has been filtered in for the highest average wind speed in a 48 hour interval, the hourly data can be used to find nails. The hourly data is combined with the merged AWS and 10 min data. Then filtering is applied on the average wind. If the average wind speed differs the rows are dropped. After filtering out nails, most stations have nails in less that 10\% measurements. The stations that have higher than 10\% error are ignored.

>>>>

This project used three main data sources, which need to be queried, filtered and combined to prepare the data for use in the models. When working with hundred of thousands of rows, the efficiency of the code is very important. Iterating through those rows might be necessary at times but will increase the time exponentially as compared to using vectorizing methods were possible. The three data sources were all in a different format. Measurement data from IMO was in text files, elevation data was in GeoTiff and reanalysis data from CARRA was in a GRIB format. To use the data to train, these three data sources needed to be combined into one file. This was done based on the measurement data from the IMO. A limit was set on the average wind speed and it was used to select measurement points. Along with the average wind speed having to be above a certain limit, to ensure that the same weather for the same location is not duplicated. The data from IMO was supplied for 10 minute increments, while CARRA data is in 3 hour intervals. This means that to use the CARRA data to predict the measured values from IMO, temporal interpolation would need to be done. Along with the temporal interpolation, note that the CARRA data is given in a rectangular grid where the distance between each point is around 2.5 km while the the information from the IMO is given at specific locations. The elevation information was given by a 20 by 20 rectangular grid that covers Iceland. When combining these data sources an interpolation method needs to be decided upon. Here linear interpolation was applied, both temporally and spatially. As the measurement points chosen were not outliers, that is the strongest average wind speed in a given range, interpolating in time does not work well. This necessitates using only measurements that are exactly at three hour intervals (00, 03, 06, 09, 12, 15, 18, 21 hours) at interpolating only in space. The choice of interpolation method, although potentially impactful on the results, was not specifically addressed in this study.

@@

Once CARRA data has been merged with AWS data, using station and time columns, then this combined file needs to checked for nails. This is done by using the hourly data (which is supposedly error free). As the data has been filtered in for the highest average wind speed in a 48 hour interval, the hourly data can be used to find nails. The hourly data is combined with the merged AWS and 10 min data. Then filtering is applied on the average wind. If the average wind speed for the 10 minute data is higher than the hourly the rows are dropped. Most stations have nails in less that 10\% measurements. The stations that have higher than 10\% error are ignored.
